---
title: "Mini-Project 01"
author: "Melanie Najera `mnajera4001@floridapoly.edu`"
date: "`r Sys.Date()`"
always_allow_html: true
output:
  md_document:
    variant: gfm
  rmdformats::robobook:
    highlight: kate
    toc_depth: 3
---

# Pre-requisites

```{r Pre-reqs,message=FALSE,warning=FALSE}
library(tidyverse)
library(dplyr)
library(gt)
library(plotly)
set.seed(636)
```

```{r Dataset}
url <- "https://raw.githubusercontent.com/mela636/dataviz_final_project/main/data/comic_books_10000_dataset.csv"
comics <- read.csv(url)
```

```{r head}
head(comics)
```
```{r glimpse}
glimpse(comics)
```

---

# Introduction
The purpose of this report is to explore the **"Comic Books Dataset (10,000 entries)"** dataset provided by *Rudra Kumar Gupta* from Kaggle. It is composed of 17 features and 10,000 observations. Certain features include the title of the comic, writer, artist (as there are situations where they are seperate from the writer), year released, genre, etc. that encompasses each comic. I wish to explore the the distribution of page counts and ratings, publishing trends over time, and how country of origin and publisher relate to the volume and reception of works within the dataset.

---

# Data Cleaning & Wrangling
To make the features more clean, they are renamed to be all lowercase and use "_" rather than periods. Additionally, create a new variable called "awarded" and "decade" in order to explore if a published work received any type of award, as well as classify that which decade the work was published in rather than by year. <br>

```{r}
comics <- comics %>%
  rename(
    comic_id = comic_id,
    title = Title,
    writer = Writer,
    artist = Artist,
    publisher = Studio.Publisher,
    year = Release.Year,
    format = Format,
    color_style = Theme..Color.Style.,
    genre = Genre,
    country = Country.of.Origin,
    pages = Page.Count,
    rating = Rating..out.of.10.,
    status = Status,
    language = Language,
    age_rating = Age.Rating,
    awards = Awards,
    volumes = Volume.Count
  )
```


**awarded**: Binary variable that marks if an award has been given.<br>
**decade**: Numerical variable that states which decade a work was published.

```{r featEng}
comics <- comics %>%
  mutate(
    awarded = awards != "None",
    decade = paste0(floor(year / 10) * 10, "s"))
```

<br> In order to make sure there are no missing values, the purpose of this chunk was to view if there were any observations with values not available. Of which, it demonstrates there are none. <br>

```{r missVar}
comics %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "column", values_to = "n_missing") %>%
  filter(n_missing > 0)
```

---

# Summary Statistcs
For the pages available, there is a minimum of 48 pages- being the shortest published work. On the other hand, there is a maximum of 14,400 pages being the longest published work. There is a large range with a clear outlier that is impacting the mean (2,232 pages) being over 600 pages more than the median. The lowest rating recorded is a score of 6 with the highest being 9.9. As for volumes, there is a minimum of 1. There is no completion status variable, however it would be interesting to see the volume count of completed works. In my personal experience, many romance manga I read are one-shots: a standalone comic that tells a complete story. As for other genres of manga, such as shounen, are longer series. It would be nice to possibly explore a relationship between volumes of completed works and genres. <br> 

```{r summary}
summary(comics[c("pages", "rating", "volumes")])
```

<p> When breaking down how many comics are available by country within this dataset, the highest is Japan with 3,575 followed by the United States with 3,150 total. There is a steep decline as the highest count is below 1,000 with only 897 comics in South Korea. It's important to note that there are combinations of countries as, for example, certain comics originate from both South Korea and Japan. </p>

```{r countryCount}
comics %>%
  count(country, sort = TRUE) %>% 
  slice_head(n = 5) %>% 
  gt() %>% 
  tab_header(title = "Number of Comics by Country") %>% 
  cols_label(country = "Country", n = "Count") %>% 
  fmt_number(columns = n, decimals = 0, use_seps = TRUE) %>% 
  data_color(
    columns = n,
    palette = "viridis") %>% 
  tab_options(
    table.font.size = px(14),
    heading.title.font.size = px(18),
    column_labels.font.weight = "bold")

```

Breaking down the unique values of country, there are combinations of countries.
```{r}
unique(comics$country)
```

<p> When breaking down the ratings by country, rather than creating a fair assessment, France/Iran and France/Spain dominate as there is only one available comic in said combination of countries, which coincidentally, have high ratings. It's unfair to compare to Japan, which the average rating of comics is 8.21 but is composed of 3,575 ratings in comparison to the one. </p>

```{r countryRate}
comics %>%
  group_by(country) %>%
  summarise(
    n = n(),
    avg_rating = round(mean(rating), 2),
    pct_awarded = round(mean(awarded) * 100, 1)
  ) %>%
  arrange(desc(avg_rating)) %>% 
  slice_head(n = 5) %>% 
  gt() %>%
  tab_header(title = "Average Ratings of Comics by Country") %>%
  cols_label(
    country = "Country",
    n = "Count",
    avg_rating = md("Average<br>Rating"),
    pct_awarded = md("Percent of<br>Works Awarded")
  ) %>%
  fmt_integer(columns = n, use_seps = TRUE) %>%
  fmt_number(columns = avg_rating, decimals = 2) %>%
  fmt_number(columns = pct_awarded, decimals = 1) %>%
  data_color(
    columns = avg_rating,
    palette = "viridis") %>%
  cols_width(
  country ~ px(160),
  n ~ px(80),
  avg_rating ~ px(110),
  pct_awarded ~ px(130)) %>% 
  tab_options(
    table.font.size = px(14),
    heading.title.font.size = px(18),
    column_labels.font.weight = "bold") 
```

<p> For the works featured that were given awards in the time of the dataset being recorded, there are 3,984 comics that were given an award- making up almost 40% of the dataset. As for the works that were not given any sort of reward, there are 6,016 comics- making up about 60% comics recorded. </p>

```{r awards}
comics %>%
  count(awarded) %>%
  mutate(
    pct = round(n / sum(n) * 100, 1),
    awarded = ifelse(awarded, "Yes", "No"),
    awarded = factor(awarded, levels = c("Yes", "No"))) %>%
  arrange(awarded) %>%
  gt() %>%
  tab_header(title = "Number of Comics Awarded") %>%
  cols_label(awarded = "Awarded", n = "Count", pct = "Percent") %>%
  fmt_number(columns = n, decimals = 0, use_seps = TRUE) %>%
  fmt_number(columns = pct, decimals = 1) %>%
  data_color(
    columns = awarded,
    palette = "viridis") %>%
  tab_options(
    table.font.size = px(14),
    heading.title.font.size = px(18),
    column_labels.font.weight = "bold")
```

---

# Visualizations

### Initial Design of: Spread of Page Count in Published Works

```{r fig.alt = "The inital design of a boxplot demonstrating the spread of page counts for 10,000 comic books. It utilizes a non-color-blind safe palette (pinks) with a lack of title, caption, and labels. It is a static image. It shows most works fall under 5,000 pages with a median of 1,571. A small number of outliers exceed 8,000 pages, with the largest reaching 14,400, pulling the mean over 600 pages above the median. However, it's difficult to tell with the lack of description."}
comics %>%
  ggplot(aes(x = pages, y = "")) +
  geom_jitter(color = "#e91e63", alpha = 0.2, height = 0.3) +
  geom_boxplot(fill = "#fce4ec", color = "#e91e63", alpha = 0.5, outlier.shape = NA) +
  scale_x_continuous(labels = scales::comma) +
  labs(x = "Pages", y = NULL) +
  theme_minimal(base_size = 13)
```

### Spread of Page Count in Published Works

```{r fig.alt = "A boxplot demonstrating the spread of page counts for 10,000 comic books. Most works fall under 5,000 pages with a median of 1,571. A small number of outliers exceed 8,000 pages, with the largest reaching 14,400, pulling the mean over 600 pages above the median.", warning=FALSE}

comics %>%
  mutate(max_val = pages == max(pages)) %>%
  ggplot(aes(x = pages, y = "")) +
    geom_jitter(color = "#440154", alpha = 0.2, height = 0.3) +
  geom_boxplot(fill = "white", color = "#440154", alpha = 0.5, outlier.shape = NA) +
  geom_label(data = ~ filter(., max_val),
             aes(label = scales::comma(pages)),
             vjust = -2, size = 3,
             color = "#440154", fill = "white",
             label.size = 0.4) +
  scale_x_continuous(labels = scales::comma) +
  annotate("label", x = 12000, y = 1.4,
           label = "Outliers pull mean 600+ pages above the median",
           color = "#440154", size = 3.5, hjust = 1,
           fill = "white", label.size = 0.5) +
  labs(title = "Spread of Page Count in Published Works",
       subtitle = "Median count of 1,571 pages",
       x = "Pages", y = NULL,
       caption = "Source: Comic Books Dataset (10,000 entries) by Rudra Kumar Gupta")+
  theme_minimal(base_size = 13)
```

### Distribution of Page Count in Published Works

```{r fig.alt = "A histogram demonstrating the spread of page counts for 10,000 comic books. Taking a closer look in comparison to the boxplot, there's a limit of 7,000 pages. Plot shows pages are right skewed with most works under 2,000 pages.", warning=FALSE}
comics %>%
  ggplot(aes(x = pages)) +
  geom_histogram(fill = "#440154", color = "white", bins = 40) +
  scale_x_continuous(labels = scales::comma, limits = c(0, 7000)) +
  annotate("label", label = "Right Skewed: Most works under 2,000 pages",
           x = 5100, y = 600, color = "#440154", size = 3.5) +
  labs(
    title = "Distribution of Page Count in Published Works",
    x = "Pages",
    y = NULL,
    caption = "Source: Comic Books Dataset (10,000 entries) by Rudra Kumar Gupta")+
  theme_minimal(base_size = 13)
```

### Volume Count of Published Works

```{r fig.alt = "A histogram demonstrating the distribution of the how many volumes are there per published works. It is right skewed with a median of 10 and mean of 13 volumes. There is a max count of 56 volumes.", warning=FALSE}
comics %>%
  ggplot(aes(x = volumes)) +
  geom_histogram(fill = "#440154", color = "white", bins = 45) +
  scale_x_continuous(limits = c(0,60))+
  annotate("label", label = "Right Skewed: Median of 10",
           x = 45, y = 600, color = "#440154", size = 3.5) +
  labs(
    title = "Volume Count of Published Works",
    x = "Volumes",
    y = NULL,
    caption = "Source: Comic Books Dataset (10,000 entries) by Rudra Kumar Gupta") +
  theme_minimal(base_size = 13)
```

### Count Distribution of Comics throughout years 2000-2026

```{r fig.alt = "A bar chart demonstrating the distribution of the how many works are publish each year from 2000-2026. Gradual increase each year- begins to dip after 2023.", message=FALSE}
years <- comics %>%
  count(year) %>%
  ggplot(aes(x = year, y = n, text = paste("Year:", year, "<br>Count:", n))) +
  geom_col(fill = "#440154") +
  coord_flip() +
  scale_x_continuous(breaks = seq(2000, 2026, by = 1)) +
  labs(title = "Count Distribution of Comics throughout Years 2000–2026 (Interactive)",
       x = NULL, y = NULL) +
  theme_minimal(base_size = 13) +
  theme(panel.grid = element_blank())

ggplotly(years, tooltip = "text") %>%
  layout(hoverlabel = list(bgcolor = "white",
                           font = list(color = "#440154")))
```

<p>You may interact with this chart by hovering over each bar. It reveals the exact count of works published that year as it's difficult to read off the graph alone. </p>

### Average Ratings of Comics Available

```{r fig.alt = "A histogram demonstrating the distribution of the average ratings of the comics that were provided. Both the median and mean are about 8.There is minimum rating score of 6 and maximum of 9.9."}
comics %>%
  ggplot(aes(x = rating)) +
  geom_histogram(fill = "#440154", color = "white", binwidth = 0.1) +
  annotate("label", x = 6.5, y = 200,
           label = "Ratings cluster between\n7.7 and 8.4 (IQR)",
           color = "#440154", size = 3.5) +
  labs(
    title = "Average Ratings of Comics Available",
    x = "Rating (out of 10)",
    y = NULL,
    caption = "Source: Comic Books Dataset (10,000 entries) by Rudra Kumar Gupta") +
  theme_minimal(base_size = 13)
```

### Average Ratings Broken Down by Country

```{r fig.alt = "A bar chart demonstrating the top 10 highest average scores broken by countries. Japan has the highest score, however there is small range of 7.9-8.2 between the listed countries."}
comics %>%
  group_by(country) %>%
  summarise(
    n = n(),
    avg_rating = mean(rating)
  ) %>%
  filter(n >= 100) %>%     
  arrange(desc(avg_rating)) %>%
  slice_head(n = 10) %>%
  mutate(country = fct_reorder(country, avg_rating)) %>%
  ggplot(aes(x = avg_rating, y = country, fill = avg_rating)) +
  geom_col() +
  scale_fill_viridis_c(option = "viridis", direction = 1) +
  scale_x_continuous(limits = c(0, 10)) +
  labs(
    title = "Top 10 Highest Average Ratings by Country (ies)",
    x = "Average Rating",
    y = NULL,
    subtitle = "Only countries with 100+ comics shown to avoid single-entry bias",
    caption = "Source: Comic Books Dataset (10,000 entries) by Rudra Kumar Gupta",
    fill = "Score") +
  theme_minimal(base_size = 13)
```

### Publishers by Works in Japan

```{r fig.alt = "A bar chart demonstrating the top ten publishers by works in Japan in descending order. Shogakukan (Viz Media) has the most published works. Everything following are close to one another. "}
comics %>%
  filter(country == "Japan") %>%
  count(publisher, sort = TRUE) %>%
  slice_head(n = 10) %>%
  mutate(publisher = fct_reorder(publisher, n)) %>%
  ggplot(aes(x = n, y = publisher)) +
  geom_col(fill = "#440154") +
  annotate("label", x = 400, y = 4,
           label = "Shogakukan (Viz Media)\n has a large lead ahead\n of its competitors.",
           color = "#440154", size = 3.2) +
  labs(
    title = "Top Ten Publishers by Works in Japan",
    x = NULL, y = NULL,
    caption = "Source: Comic Books Dataset (10,000 entries) by Rudra Kumar Gupta") +
  theme_minimal(base_size = 13) +
  theme(panel.grid = element_blank())
```

### Page Count by Publisher in Japan

```{r fig.alt = "A boxplot chart demonstrating the page counts seperated by the publishers in Japan. "}
comics %>%
  filter(country == "Japan") %>%
  mutate(publisher = str_remove(publisher, "\\s*\\(.*\\)")) %>%
  count(publisher) %>%
  filter(n >= 50) %>%
  inner_join(
    comics %>%
      filter(country == "Japan") %>%
      mutate(publisher = str_remove(publisher, "\\s*\\(.*\\)")),
    by = "publisher"
  ) %>%
  mutate(publisher = fct_reorder(publisher, pages, median)) %>%
  ggplot(aes(x = pages, y = publisher, color = publisher, fill = publisher)) +
  geom_boxplot(outlier.alpha = 0.4) +
  scale_color_viridis_d(option = "viridis") +
  scale_fill_viridis_d(option = "viridis", alpha = 0.3) +
  scale_x_continuous(labels = scales::comma) +
  annotate("label", x = 12000, y = 6.5,
           label = "Wide spread within publishers\nsuggests varied series lengths",
           color = "#440154", size = 3.2)+
  labs(
    title = "Page Count by Publisher in Japan",
    x = "Pages", y = NULL,
    caption = "Source: Comic Books Dataset (10,000 entries) by Rudra Kumar Gupta") +
  theme_minimal(base_size = 13) +
  theme(panel.grid = element_blank(), legend.position = "none")
```

---

# Project Questions
## What were the original charts you planned to create for this assignments?
I original wanted to explore more about the genres that are provided- particularly a bar chart that would demonstrate the top ten most popular genres within the dataset. As well as to explore the relationship between manga genres and their volume count and see if it would reflect my personal experience of which genres have more one-shots compared to longer running series. Prior to exploring the variables provided, I wanted to explore completed works to see their final totals of ratings to reflect the entire comic; this is more so for works that plummet or skyrocket in ratings depending on the ending. Additionally, I'd hoped that pages would be for per volume as, majority of manga volumes I see print the same amount of pages depending on the publishing company; I wanted to see if the data would align with that belief. <br>

## What story could you tell with your plots?
The story told with the plots is of one that the comic landscape is dominated by Japanese and American publishers with clear differences in scale and reach. The page count distribution is heavily right-skewed, with most works falling below 3,500 pages but a small number of long works pulling the mean well above the median, where most series are shorter runs and few sustain a large number of volumes. The year distribution reveals that comic publishing in this dataset building up and peaked in 2023 before falling slightly. The rating distribution shows that the majority of comics score between 7.5 and 9.0, which may suggest that there's either an actual well-received collection or the dataset leaned toward gathering notable works rather than obscure ones. Which, it's important to note that the lowest rating provided is a 6. When breaking down ratings by country, South Korea and Japan score the highest, however it's a very gradual decrease in ratings. Looking into Japan, Shogakukan (Viz Media) dominates in sheer number of titles published, although page count distributions across Japanese publishers are similar, suggesting that total pages is more a function of how many volumes a series runs rather than which house publishes it. <br>

## How did you apply the principles of data visualizations and design for this assignment?
In applying the principles of data visualizations and design, I originally applied a consistent pink color palette (#fce4ec to #e91e63) for all of the tables and plots to create visual cohesion throughout the report. Currently, I utilize the viridis palette. To keep the format of the html simple while adding a bit of style, I use the robobook rmdformat. I removed certain chart elements such as certain grid lines to keep the background uncrowded and minimized elements with theme_minimal(). For the tables, I highlighted the counts with color gradients to direct the focus of the viewer. I scaled the x-axis to prevent extreme outliers shrinking where the majority of data grouped. I filtered graphs to have a minimum amount of comics per country to not have an unfair comparison between single-entry country combinations to established ones. To avoid being repetitive with the graphs, I titled them in the headers and did not put them again in the graph itself to keep it less cluttered. Currently, there are titles if there is a need to separate the figure from the repeat. Lastly, if the features were long in name, to flip the axis to avoid having diagonal names to prevent the viewer from tilting their head. 

## Redesigns of Charts
The initial color palette of this project was pink (#fce4ec to #e91e63) for all of the tables and plots to create visual cohesion throughout the report, however, each chart now follows the viridis palette as it's **color-blind safe**. Additionally, each now have a title to not rely on the report's section header and may be read on its own. There are labels- such as "Outliers pull mean 600+ pages above the median" to help point out the impact the graph is demonstrating, as well as point out the points on the graph that stand out. Each have an alternative description to serve as an aid to the audience. The "Count Distribution of Comics throughout years 2000-2026" was altered to be interactive for when the user hovers, it will display how many works were published in said year. I added a proper palette to "Page Count by Publisher In Japan so each box is no longer the same color.
